---
title: "Emma Pan Neuro (Control + ND75KD) - Preprocessing and plotting GSE243639 from Martirosyan et al."
author: "Vincent Gardeux"
date: "2025/02/27"
last_modified: "2025/03/04"
output:
  html_document:
    df_print: paged
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/data/gardeux/Neuro_Droso_ND75KD/")
```

## Libraries & functions

First, I'm loading the required libraries & functions

```{r}
suppressPackageStartupMessages(library(Seurat, lib.loc = "/software/Seuratv4/lib/")) # Seurat v4.4.0, library for single-cell analysis
suppressPackageStartupMessages(library(SeuratObject, lib.loc = "/software/Seuratv4/lib/")) # For compatibility purpose
suppressPackageStartupMessages(library(data.table)) # For writing DE gene file
suppressPackageStartupMessages(library(readxl)) # For writing DE gene file
suppressPackageStartupMessages(library(loomR)) # For building a file for ASAP
suppressPackageStartupMessages(library(Matrix)) # For building a file for ASAP
suppressPackageStartupMessages(library(ggplot2)) # For plotting
suppressPackageStartupMessages(library(ggpubr)) # For grid plotting
suppressPackageStartupMessages(library(plotly)) # For interactive plots
suppressPackageStartupMessages(library(crayon)) # Just for bolding the console output :D

cat(bold("Seurat"), "version", as.character(packageVersion("Seurat")), "\n")
cat(bold("SeuratObject"), "version", as.character(packageVersion("SeuratObject")), "\n")
cat(bold("data.table"), "version", as.character(packageVersion("data.table")), "\n")
cat(bold("readxl"), "version", as.character(packageVersion("readxl")), "\n")
cat(bold("loomR"), "version", as.character(packageVersion("loomR")), "\n")
cat(bold("Matrix"), "version", as.character(packageVersion("Matrix")), "\n")
cat(bold("ggplot2"), "version", as.character(packageVersion("ggplot2")), "\n")
cat(bold("ggpubr"), "version", as.character(packageVersion("ggpubr")), "\n")
cat(bold("plotly"), "version", as.character(packageVersion("plotly")), "\n")

# Color palettes
steinPalette <- c(`Central Brain A`="#80C980",`Central Brain B`="#BDADD4",`Optic lobe`="#376CB0",`Glia`="#FBBF85",`Other`="#F0027E")

# Random seed
set.seed(42)
```

## Parameters

*Paper:*
Unravelling cell type-specific responses to Parkinson’s Disease at single cell resolution
Martirosyan et al.
Mol Neurodegeneration 19, 7 (2024)
https://molecularneurodegeneration.biomedcentral.com/articles/10.1186/s13024-023-00699-0

*Data repo:* https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE243639

*Loom file:* NA

```{r}
## Parameters
input_matrix <- "./data/Martirosyan_2024/GSE243639_Filtered_count_table.csv.gz"
input_metadata <- "./data/Martirosyan_2024/GSE243639_UMAP_coordinates.xlsx"
input_clinical <- "./data/Martirosyan_2024/GSE243639_Clinical_data.csv"
input_Oxphos_135_AUCell_values_path <- "./data/Martirosyan_135_Oxphos_AUCell_auc.tsv"
input_UPR_93_AUCell_values_path <- "./data/Martirosyan_93_UPR_AUCell_auc.tsv"
input_SCENIC_regulons_path <- "./data/Martirosyan_2024/Martirosyan_regulons_aucell.tsv"
input_SCENIC_regulons_binarized_path <- "./data/Martirosyan_2024/Martirosyan_regulons_aucell_binarized.tsv"

output_seurat_final <- "./data/Martirosyan_2024/Martirosyan"
output_de_folder <- "./data/Martirosyan_2024/"
```

## I. Loading dataset

### I.1. Reading count matrix from csv provided by authors on GEO

```{r}
data.matrix <- fread(input_matrix, header = T, stringsAsFactors = F, showProgress = F, sep = ",", data.table = F, check.names = F)
rownames(data.matrix) <- data.matrix$V1
data.matrix <- data.matrix[,-1]
```

### I.2. Build Seurat object

```{r}
# Build a Seurat object with this data
data.seurat <- CreateSeuratObject(data.matrix, project = "Martirosyan")

# Restore hyphens in cell barcodes
data.seurat <- RenameCells(data.seurat, new.names = gsub(pattern = ".1$", replacement = "-1", x = colnames(data.seurat)))

# Change identity
data.seurat$orig.ident.parsed <- data.seurat$orig.ident
data.seurat$orig.ident <- "Martirosyan"
data.seurat <- SetIdent(data.seurat, value = "orig.ident")

# Add MT content
data.seurat[["percent.mt"]] <- PercentageFeatureSet(data.seurat, pattern = "^MT-")

# Run default Normalization
data.seurat <- NormalizeData(data.seurat, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
# Visualize QC metrics as a violin plot
VlnPlot(data.seurat, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

### I.3. Reading global metadata from authors

```{r}
data.metadata <- as.data.frame(read_xlsx(path = input_metadata, sheet = 1))
rownames(data.metadata) <- data.metadata$CELL_ID
data.metadata$UMAP_1 <- as.numeric(data.metadata$UMAP_1)
data.metadata$UMAP_2 <- as.numeric(data.metadata$UMAP_2)
data.metadata
```

## I.4 Reading OXPHOS AUC values from AUCell (using the 133 OXPHOS genes from KEGG Oxphos pathway https://www.genome.jp/entry/mmu00190)

```{r}
data.auc.oxphos <- fread(input_Oxphos_135_AUCell_values_path, data.table = F, check.names = F)
rownames(data.auc.oxphos) <- data.auc.oxphos$Cell
data.auc.oxphos
```


## I.5  Reading UPR AUC values from AUCell (using the 93 UPR genes from REACTOME

```{r}
data.auc.upr <- fread(input_UPR_93_AUCell_values_path, data.table = F, check.names = F)
rownames(data.auc.upr) <- data.auc.upr$Cell
data.auc.upr
```

## I.6 Add to Seurat Object

```{r}
# Add metadata
data.seurat$AUC_OXPHOS_135_KEGG <- data.auc.oxphos[colnames(data.seurat), "AUC"]
data.seurat$AUC_UPR_93_REACTOME <- data.auc.upr[colnames(data.seurat), "AUC"]
data.seurat$CLUSTER <- data.metadata[colnames(data.seurat), "CLUSTER"]
data.seurat$IDENT <- data.metadata[colnames(data.seurat), "IDENT"]

# Add UMAP
data.seurat[["UMAP"]] <- CreateDimReducObject(embeddings = as.matrix(data.metadata[,c("UMAP_1", "UMAP_2")]), key = "UMAP_", assay = "RNA")
```

## II Plot interesting stuff

### II.1 Plot author's UMAP

```{r}
DimPlot(data.seurat, reduction = "UMAP", group.by = "IDENT")
```

```{r}
DimPlot(data.seurat, reduction = "UMAP", group.by = "CLUSTER")
```

### II.2.a Plot Distribution OXPHOS

```{r}
hist(data.seurat$AUC_OXPHOS_135_KEGG, breaks = 100)
```

### II.2.b Plot Boxplot OXPHOS

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "AUC_OXPHOS_135_KEGG" for each group
median_values <- tapply(boxplot_data$AUC_OXPHOS_135_KEGG, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = AUC_OXPHOS_135_KEGG, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "AUC_OXPHOS_135_KEGG") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

### II.2.c Plot UMAP OXPHOS

```{r}
FeaturePlot(data.seurat, reduction = "UMAP", features = "AUC_OXPHOS_135_KEGG")
```

### II.3.a Plot Distribution UPR

```{r}
hist(data.seurat$AUC_UPR_93_REACTOME, breaks = 100)
```

### II.3.b Plot Boxplot UPR

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "AUC_UPR_93_REACTOME" for each group
median_values <- tapply(boxplot_data$AUC_UPR_93_REACTOME, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = AUC_UPR_93_REACTOME, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "AUC_UPR_93_REACTOME") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

### II.3.c Plot UMAP UPR

```{r}
FeaturePlot(data.seurat, reduction = "UMAP", features = "AUC_UPR_93_REACTOME")
```

### II.4.a Plot Boxplot HSP90AA1

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data
boxplot_data$HSP90AA1 <- data.seurat@assays$RNA@data["HSP90AA1",]

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "HSP90AA1" for each group
median_values <- tapply(boxplot_data$HSP90AA1, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = HSP90AA1, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "HSP90AA1 Normalized expression") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

### II.4.b Plot UMAP HSP90AA1

```{r}
FeaturePlot(data.seurat, reduction = "UMAP", features = "HSP90AA1")
```

### II.5.a Plot Boxplot HSPA8

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data
boxplot_data$HSPA8 <- data.seurat@assays$RNA@data["HSPA8",]

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "HSPA8" for each group
median_values <- tapply(boxplot_data$HSPA8, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = HSPA8, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "HSPA8 Normalized expression") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

### II.5.b Plot UMAP HSPA8

```{r}
FeaturePlot(data.seurat, reduction = "UMAP", features = "HSPA8")
```

### II.6.a Plot Boxplot ATF4

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data
boxplot_data$ATF4 <- data.seurat@assays$RNA@data["ATF4",]

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "ATF4" for each group
median_values <- tapply(boxplot_data$ATF4, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = ATF4, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "ATF4 Normalized expression") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

### II.6.b Plot UMAP ATF4

```{r}
FeaturePlot(data.seurat, reduction = "UMAP", features = "ATF4")
```

### III Regulon data from pySCENIC

## III.1 Reading regulons values from SCENIC output

```{r}
data.scenic <- fread(input_SCENIC_regulons_path, data.table = F, check.names = F)
rownames(data.scenic) <- gsub(x = data.scenic$Cell, pattern = ".1$", replacement = "-1")
data.scenic <- data.scenic[,-1]
data.scenic
```

Updating colnames

```{r}
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\+\\)", replacement = "_activating_regulon")
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\-\\)", replacement = "_repressing_regulon")
colnames(data.scenic) <- paste0("SCENIC_", colnames(data.scenic))
data.scenic
```

## III.2 Adding each regulon as a metadata to the Seurat object:

```{r}
# Adding each column as a metadata
for(regulon in colnames(data.scenic)){
  data.seurat@meta.data[[regulon]] <- data.scenic[colnames(data.seurat), regulon]
}
```

## III.3 Compute UMAP on regulon matrix

```{r, fig.height = 8, fig.width = 12}
# Adding whole regulon matrix as an embedding
# Saving mapping for later
data.mapping.regulon <- data.frame(regulon = colnames(data.scenic), mapping = paste0("SCENIC_", 1:ncol(data.scenic)))
rownames(data.mapping.regulon) <- data.mapping.regulon$regulon
# Change col names to comply with required format
colnames(data.scenic) <- data.mapping.regulon[colnames(data.scenic), "mapping"]
data.seurat@reductions[['SCENIC']] <- CreateDimReducObject(embeddings = as.matrix(data.scenic), key = 'SCENIC_', assay = 'RNA')

# Compute UMAP on it
data.seurat <- RunUMAP(object = data.seurat, reduction.name = "SCENICumap", reduction = "SCENIC", dims = 1:ncol(data.scenic), reduction.key = "SCENICumap_", assay = 'RNA', n.neighbors = 10L, metric = "correlation", min.dist = 0.4)
DimPlot(data.seurat, reduction = "SCENICumap", label = TRUE, pt.size = 0.5, group.by = "IDENT") + NoLegend()
```

### IV Binarized Regulon data from pySCENIC

## IV.1 Reading regulons values from binarized SCENIC output

```{r}
data.scenic <- fread(input_SCENIC_regulons_binarized_path, data.table = F)
rownames(data.scenic) <- gsub(x = data.scenic$Cell, pattern = ".1$", replacement = "-1")
data.scenic <- data.scenic[,-1]
data.scenic
```

Updating colnames

```{r}
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\+\\)", replacement = "_activating_regulon_binarized")
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\-\\)", replacement = "_repressing_regulon_binarized")
colnames(data.scenic) <- paste0("SCENIC_", colnames(data.scenic))
data.scenic
```

## IV.2 Adding each regulon as a metadata to the Seurat object:

```{r}
# Adding each column as a metadata
for(regulon in colnames(data.scenic)){
  data.seurat@meta.data[[regulon]] <- data.scenic[colnames(data.seurat), regulon]
}
```

## IV.3 Compute UMAP on binarized regulon matrix

```{r, fig.height = 8, fig.width = 12}
# Adding whole binarized regulon matrix as an embedding
# Saving mapping for later
data.mapping.regulon_binarized <- data.frame(regulon = colnames(data.scenic), mapping = paste0("SCENIC_", 1:ncol(data.scenic)))
rownames(data.mapping.regulon_binarized) <- data.mapping.regulon_binarized$regulon
# Change col names to comply with required format
colnames(data.scenic) <- data.mapping.regulon_binarized[colnames(data.scenic), "mapping"]
data.seurat@reductions[['SCENICbinarized']] <- CreateDimReducObject(embeddings = as.matrix(data.scenic), key = 'SCENICbinarized_', assay = 'RNA')

# Compute UMAP on it
data.seurat <- RunUMAP(object = data.seurat, reduction.name = "SCENICbinarizedumap", reduction = "SCENICbinarized", dims = 1:ncol(data.scenic), reduction.key = "SCENICbinarizedumap_", assay = 'RNA', n.neighbors = 10L, metric = "correlation", min.dist = 0.4)
DimPlot(data.seurat, reduction = "SCENICbinarizedumap", label = TRUE, pt.size = 0.5, group.by = "IDENT") + NoLegend()
```

## IV.4 ATF4 regulon

```{r, fig.width=10, fig.height=5}
FeaturePlot(data.seurat, reduction = "UMAP", label = F, pt.size = 0.5, features = c("SCENIC_ATF4_activating_regulon", "SCENIC_ATF4_activating_regulon_binarized")) + NoLegend()
```

```{r, fig.width=10, fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICumap", label = F, pt.size = 0.5, features = c("SCENIC_ATF4_activating_regulon", "SCENIC_ATF4_activating_regulon_binarized")) + NoLegend()
```

```{r, fig.width=10, fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", label = F, pt.size = 0.5, features = c("SCENIC_ATF4_activating_regulon", "SCENIC_ATF4_activating_regulon_binarized")) + NoLegend()
```

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "ATF4 regulon" for each group
median_values <- tapply(boxplot_data$SCENIC_ATF4_activating_regulon, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = SCENIC_ATF4_activating_regulon, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "ATF4 regulon activity") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

## IV.5 Clinical data

## IV.5.1 Read clinical data

```{r}
data.clinical <- fread(input_clinical, sep = ";", header = T, skip = 5, data.table = F)
data.clinical
```

### IV.5.2 Adding to Seurat object

```{r}
data.seurat$clinicalDiag <- data.clinical$`Clinical diagnosis`[match(data.seurat$orig.ident.parsed, data.clinical$`Sample ID`)]
data.seurat$clinicalDiag <- factor(data.seurat$clinicalDiag, levels = c("Control", "Parkinson's"))
data.seurat$sex <- data.clinical$Sex[match(data.seurat$orig.ident.parsed, data.clinical$`Sample ID`)]
data.seurat$sex <- factor(data.seurat$sex, levels = c("female", "male"))
data.seurat$age <- data.clinical$Age[match(data.seurat$orig.ident.parsed, data.clinical$`Sample ID`)]
data.seurat$pmi <- data.clinical$`PMI hours`[match(data.seurat$orig.ident.parsed, data.clinical$`Sample ID`)]
data.seurat$lewy_midbrain <- data.clinical$`Lewy bodies presence in midbrain`[match(data.seurat$orig.ident.parsed, data.clinical$`Sample ID`)]
data.seurat$lewy_midbrain[is.na(data.seurat$lewy_midbrain)] <- 0
data.seurat$lewy_amygdala <- data.clinical$`Lewy bodies presence in limbic regions (amygdala)`[match(data.seurat$orig.ident.parsed, data.clinical$`Sample ID`)]
data.seurat$lewy_amygdala[is.na(data.seurat$lewy_amygdala)] <- 0
data.seurat$lewy_cortex <- data.clinical$`Lewy bodies presence in neocortical regions (frontal cortex)`[match(data.seurat$orig.ident.parsed, data.clinical$`Sample ID`)]
data.seurat$lewy_cortex[is.na(data.seurat$lewy_cortex)] <- 0
data.seurat$cerad_score_plaque <- data.clinical$`CERAD score for neuritic plaques`[match(data.seurat$orig.ident.parsed, data.clinical$`Sample ID`)]
data.seurat$cerad_score_plaque[is.na(data.seurat$cerad_score_plaque)] <- 0
data.seurat$braak_stage <- data.clinical$`Braak stage for neurofibrillary tangles`[match(data.seurat$orig.ident.parsed, data.clinical$`Sample ID`)]
data.seurat$braak_stage[is.na(data.seurat$braak_stage)] <- 0
```

## IV.6 Save Seurat object

```{r}
saveRDS(data.seurat, file = paste0(output_seurat_final, ".rds"))
```

## V Subsetting to only neurons

### V.1 Read metadata from authors

```{r}
data.neurons <- as.data.frame(read_xlsx(path = input_metadata, sheet = "Neurons"))
rownames(data.neurons) <- data.neurons$CELL_ID
data.neurons$UMAP_1 <- as.numeric(data.neurons$UMAP_1)
data.neurons$UMAP_2 <- as.numeric(data.neurons$UMAP_2)
data.neurons
```

Authors have subset their dataset and redo a UMAP and clustering on the subset dataset.

### V.2 Subset the Seurat object

```{r}
# Subset to neuron cells from metadata
neuron.cells <- rownames(data.neurons)[rownames(data.neurons) %in% colnames(data.seurat)]
data.seurat <- subset(data.seurat, cells = neuron.cells)

# Add/replace new metadata
data.seurat$CLUSTER <- data.neurons[colnames(data.seurat), "CLUSTER"]
data.seurat$IDENT <- data.neurons[colnames(data.seurat), "IDENT"]

# Add UMAP
data.seurat[["UMAP"]] <- CreateDimReducObject(embeddings = as.matrix(data.neurons[,c("UMAP_1", "UMAP_2")]), key = "UMAP_", assay = "RNA")
```

### V.3 Recompute regulon UMAPs from SCENIC vectors

#### V.3.1 Compute UMAP on regulon matrix

```{r, fig.height = 5, fig.width = 5}
# Capture the non-binarized regulon matrix from metadata
data.scenic <- data.seurat@meta.data
data.scenic <- data.scenic[,endsWith(colnames(data.scenic), suffix = "_activating_regulon")]

# Adding whole regulon matrix as an embedding
colnames(data.scenic) <- paste0("SCENIC_", 1:ncol(data.scenic))
data.seurat@reductions[['SCENIC']] <- CreateDimReducObject(embeddings = as.matrix(data.scenic), key = 'SCENIC_', assay = 'RNA')

# Compute UMAP on it
data.seurat <- RunUMAP(object = data.seurat, reduction.name = "SCENICumap", reduction = "SCENIC", dims = 1:ncol(data.scenic), reduction.key = "SCENICumap_", assay = 'RNA', n.neighbors = 10L, metric = "correlation", min.dist = 0.4)
DimPlot(data.seurat, reduction = "SCENICumap", label = TRUE, pt.size = 0.5, group.by = "IDENT") + NoLegend()
```

#### V.3.2 Compute UMAP on binarized regulon matrix

```{r, fig.height = 5, fig.width = 5}
# Capture the binarized regulon matrix from metadata
data.scenic <- data.seurat@meta.data
data.scenic <- data.scenic[,endsWith(colnames(data.scenic), suffix = "_activating_regulon_binarized")]

# Adding whole regulon matrix as an embedding
colnames(data.scenic) <- paste0("SCENIC_", 1:ncol(data.scenic))
data.seurat@reductions[['SCENICbinarized']] <- CreateDimReducObject(embeddings = as.matrix(data.scenic), key = 'SCENICbinarized_', assay = 'RNA')

# Compute UMAP on it
data.seurat <- RunUMAP(object = data.seurat, reduction.name = "SCENICbinarizedumap", reduction = "SCENICbinarized", dims = 1:ncol(data.scenic), reduction.key = "SCENICbinarizedumap_", assay = 'RNA', n.neighbors = 10L, metric = "correlation", min.dist = 0.4)
DimPlot(data.seurat, reduction = "SCENICbinarizedumap", label = TRUE, pt.size = 0.5, group.by = "IDENT") + NoLegend()
```

## VI Plot interesting stuff

### VI.1 Plot author's UMAP

```{r, fig.height = 5, fig.width = 5}
DimPlot(data.seurat, reduction = "UMAP", group.by = "IDENT", label = T) + NoLegend()
```

```{r, fig.height = 5, fig.width = 5}
DimPlot(data.seurat, reduction = "UMAP", group.by = "CLUSTER", label = T) + NoLegend()
```

### VI.2 OXPHOS score(gene set from KEGG, AUCell for scoring)

#### VI.2.a Plot Distribution OXPHOS

```{r}
hist(data.seurat$AUC_OXPHOS_135_KEGG, breaks = 100)
```

#### VI.2.b Plot Boxplot OXPHOS

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "AUC_OXPHOS_135_KEGG" for each group
median_values <- tapply(boxplot_data$AUC_OXPHOS_135_KEGG, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = AUC_OXPHOS_135_KEGG, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "AUC_OXPHOS_135_KEGG") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

#### VI.2.c Plot UMAP OXPHOS

```{r, fig.height = 5, fig.width = 6}
FeaturePlot(data.seurat, reduction = "UMAP", features = "AUC_OXPHOS_135_KEGG")
```

```{r, fig.height = 5, fig.width = 6}
FeaturePlot(data.seurat, reduction = "SCENICumap", features = "AUC_OXPHOS_135_KEGG")
```

```{r, fig.height = 5, fig.width = 6}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = "AUC_OXPHOS_135_KEGG")
```

### VI.3 UPR scoring (gene set from REACTOME, scoring with AUCell)

#### VI.3.a Plot Distribution UPR

```{r}
hist(data.seurat$AUC_UPR_93_REACTOME, breaks = 100)
```

#### IV.3.b Plot Boxplot UPR

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "AUC_UPR_93_REACTOME" for each group
median_values <- tapply(boxplot_data$AUC_UPR_93_REACTOME, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = AUC_UPR_93_REACTOME, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "AUC_UPR_93_REACTOME") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

#### VI.3.c Plot UMAP UPR

```{r, fig.height = 5, fig.width = 6}
FeaturePlot(data.seurat, reduction = "UMAP", features = "AUC_UPR_93_REACTOME")
```

```{r, fig.height = 5, fig.width = 6}
FeaturePlot(data.seurat, reduction = "SCENICumap", features = "AUC_UPR_93_REACTOME")
```

```{r, fig.height = 5, fig.width = 6}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = "AUC_UPR_93_REACTOME")
```

### VI.4 HSP90AA1 gene of interest

#### VI.4.a Plot Boxplot HSP90AA1

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data
boxplot_data$HSP90AA1 <- data.seurat@assays$RNA@data["HSP90AA1",]

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "HSP90AA1" for each group
median_values <- tapply(boxplot_data$HSP90AA1, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = HSP90AA1, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "HSP90AA1 Normalized expression") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

#### VI.4.b Plot UMAP HSP90AA1

```{r, fig.width=6,fig.height=5}
FeaturePlot(data.seurat, reduction = "UMAP", features = "HSP90AA1")
```

```{r, fig.width=6,fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICumap", features = "HSP90AA1")
```

```{r, fig.width=6,fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = "HSP90AA1")
```

### VI.5 HSPA8 gene of interest

#### VI.5.a Plot Boxplot HSPA8

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data
boxplot_data$HSPA8 <- data.seurat@assays$RNA@data["HSPA8",]

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "HSPA8" for each group
median_values <- tapply(boxplot_data$HSPA8, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = HSPA8, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "HSPA8 Normalized expression") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

#### VI.5.b Plot UMAP HSPA8

```{r, fig.width=6,fig.height=5}
FeaturePlot(data.seurat, reduction = "UMAP", features = "HSPA8")
```

```{r, fig.width=6,fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICumap", features = "HSPA8")
```

```{r, fig.width=6,fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = "HSPA8")
```

### VI.6 ATF4 gene/TF of interest

#### VI.6.a Plot Boxplot ATF4

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data
boxplot_data$ATF4 <- data.seurat@assays$RNA@data["ATF4",]

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "ATF4" for each group
median_values <- tapply(boxplot_data$ATF4, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = ATF4, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "ATF4 Normalized expression") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

#### VI.6.b Plot UMAP ATF4

```{r, fig.width=6,fig.height=5}
FeaturePlot(data.seurat, reduction = "UMAP", features = "ATF4")
```

```{r, fig.width=6,fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICumap", features = "ATF4")
```

```{r, fig.width=6,fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = "ATF4")
```

### VI.7 ATF4 regulon

#### VI.7.a Plot Boxplot ATF4 regulon activity

```{r, fig.width=5,fig.height=5}
boxplot_data <- data.seurat@meta.data

# Filter out <5 cells per annotation
boxplot_data <- boxplot_data[boxplot_data$IDENT %in% names(which(table(boxplot_data$IDENT) >= 5)),]

# Calculating median values of "ATF4" for each group
median_values <- tapply(boxplot_data$SCENIC_ATF4_activating_regulon, boxplot_data$IDENT, median)

# Sorting the metadata based on median values
boxplot_data$IDENT <- factor(boxplot_data$IDENT, levels = names(sort(median_values, decreasing = F)))

# Plot
p <- ggplot(boxplot_data, aes(x = IDENT, y = SCENIC_ATF4_activating_regulon, fill = IDENT)) +
  geom_boxplot() +
  labs(x = NULL, y = "ATF4 regulon activity") +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  NoLegend()

p
```

#### VI.6.b Plot UMAP ATF4 regulon activity

```{r, fig.width=11,fig.height=5}
FeaturePlot(data.seurat, reduction = "UMAP", features = c("SCENIC_ATF4_activating_regulon","SCENIC_ATF4_activating_regulon_binarized"))
```

```{r, fig.width=11,fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICumap", features = c("SCENIC_ATF4_activating_regulon","SCENIC_ATF4_activating_regulon_binarized"))
```

```{r, fig.width=11,fig.height=5}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("SCENIC_ATF4_activating_regulon","SCENIC_ATF4_activating_regulon_binarized"))
```

#### VII Differential gene / regulon in Neurons00 && Neurons03

#### VII.1 DE genes in Neurons00+Neurons03 vs others

```{r}
# 1. Create new grouping
data.seurat$UPC_Oxphos_Neurons <- data.seurat$IDENT %in% c("neurons00", "neurons03")
data.seurat$UPC_Oxphos_Neurons[data.seurat$UPC_Oxphos_Neurons] <- "Neurons00+Neurons03"
data.seurat$UPC_Oxphos_Neurons[data.seurat$UPC_Oxphos_Neurons == "FALSE"] <- "Others"

# 2. Run DE
gene.markers <- FindMarkers(data.seurat, assay = "RNA", slot = "data", ident.1 = "Neurons00+Neurons03", ident.2 = "Others", group.by = "UPC_Oxphos_Neurons", verbose = T, test.use = "wilcox", random.seed = 42, logfc.threshold = 0, min.pct = -Inf)

# 3. Enhance output
gene.markers$gene <- rownames(gene.markers)

# 4. Save in file
fwrite(gene.markers, file = paste0(output_de_folder, "de_genes.tsv"), sep = "\t")

# Filtered output
gene.markers.filtered <- gene.markers[gene.markers$p_val_adj <= 0.05,]
gene.markers.filtered <- gene.markers.filtered[order(abs(gene.markers.filtered$avg_log2FC), decreasing = T),]
fwrite(gene.markers.filtered, file = paste0(output_de_folder, "de_genes_filtered.tsv"), sep = "\t")

gene.markers.filtered
```

```{r, fig.width=11,fig.height=10}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("ENO1","KAZN","ATP6V0B","CSMD1","GALNTL6","HSP90AA1", "NEFM", "ALDH1A1", "TUBB2A"), ncol = 3)
```

#### VII.2 Differentially activated regulons in Neurons00+Neurons03 vs others (based on non-binarized data)

```{r}
# 1. Run DE
regulon.markers <- FindMarkers(data.seurat, assay = "SCENIC", slot = "data", ident.1 = "Neurons00+Neurons03", ident.2 = "Others", group.by = "UPC_Oxphos_Neurons", verbose = T, test.use = "wilcox", random.seed = 42, logfc.threshold = 0, min.pct = -Inf)

# 2. Enhance output
rownames(data.mapping.regulon) <- data.mapping.regulon$mapping
regulon.markers$regulon <- data.mapping.regulon[rownames(regulon.markers), "regulon"]
regulon.markers$gene <- gsub(x = regulon.markers$regulon, pattern = "SCENIC_", replacement = "")
regulon.markers$gene <- gsub(x = regulon.markers$gene, pattern = "_activating_regulon", replacement = "")

# 3. Save in file
fwrite(regulon.markers, file = paste0(output_de_folder, "de_regulons.tsv"), sep = "\t")

# Filtered output
regulon.markers.filtered <- regulon.markers[regulon.markers$p_val_adj <= 0.05,]
regulon.markers.filtered <- regulon.markers.filtered[order(abs(regulon.markers.filtered$avg_diff), decreasing = T),]
fwrite(regulon.markers.filtered, file = paste0(output_de_folder, "de_regulons_filtered.tsv"), sep = "\t")

regulon.markers.filtered
```

```{r, fig.width=16,fig.height=15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("SCENIC_ENO1_activating_regulon","SCENIC_CYB5R1_activating_regulon","SCENIC_GOT1_activating_regulon","SCENIC_PKM_activating_regulon","SCENIC_TCEAL2_activating_regulon","SCENIC_FOXA1_activating_regulon", "SCENIC_TPI1_activating_regulon", "SCENIC_TCEAL6_activating_regulon", "SCENIC_SOD1_activating_regulon"), ncol = 3)
```

#### VII.3 Save Seurat object

```{r}
saveRDS(data.seurat, file = paste0(output_seurat_final, "_neurons.rds"))
```
