---
title: "Emma Pan Neuro (Control + ND75KD) - Preprocessing Human Brain Aging dataset from Jeffries et al."
author: "Vincent Gardeux"
date: "2025/11/13"
last_modified: "2025/11/13"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/data/gardeux/Neuro_Droso_ND75KD/")
```

## Libraries & functions

First, I'm loading the required libraries & functions

```{r}
suppressPackageStartupMessages(library(Seurat)) # Seurat v5, for compatibility with .rds object
suppressPackageStartupMessages(library(SeuratObject)) # Seurat v5 object
suppressPackageStartupMessages(library(data.table)) # For writing DE gene file
suppressPackageStartupMessages(library(reticulate)) # For Python access
suppressPackageStartupMessages(library(Matrix)) # For writing DE gene file
suppressPackageStartupMessages(library(MatrixExtra)) # For writing DE gene file
suppressPackageStartupMessages(library(crayon)) # Just for bolding the console output :D

cat(bold("Seurat"), "version", as.character(packageVersion("Seurat")), "\n")
cat(bold("SeuratObject"), "version", as.character(packageVersion("SeuratObject")), "\n")
cat(bold("data.table"), "version", as.character(packageVersion("data.table")), "\n")
cat(bold("reticulate"), "version", as.character(packageVersion("reticulate")), "\n")
cat(bold("Matrix"), "version", as.character(packageVersion("Matrix")), "\n")
cat(bold("MatrixExtra"), "version", as.character(packageVersion("MatrixExtra")), "\n")

# Random seed
set.seed(42)
```

## Set Python environment (reticulate)

```{r}
reticulate::use_python("/usr/bin/python3.12", required = TRUE)
py_config()
ad <- reticulate::import("anndata") # Import anndata Python package

cat(bold("anndata"), "version", as.character(ad$`__version__`), "\n")
```

## Parameters

*Paper:*
Jeffries, A.M., Yu, T., Ziegenfuss, J.S. et al. Single-cell transcriptomic and genomic changes in the ageing human brain. Nature 646, 657–666 (2025). https://doi.org/10.1038/s41586-025-09435-8

*Data repo:* https://publications.wenglab.org/SomaMut/Jeffries_Yu_BrainAging_2025/

*Loom / H5ad file:* NA. Made in this script from .rds file.

```{r}
## Parameters
input_rds <- "./data/Jeffries_2025_HumanBrainAging/pfc.clean.rds" # 367'317 cells vs 29'729 genes
input_metadata <- "./data/Jeffries_2025_HumanBrainAging/pfc.clean.filtered_metadata.txt" # Just to check

output_h5ad <- "./data/Jeffries_2025_HumanBrainAging/pfc.clean.h5ad"

message("1. Seurat object to convert (input RDS file) = ", input_rds)
message("2. H5ad file path = ", output_h5ad)
```

## I. Loading dataset

### I.1. Reading rds file (Seurat object) downloaded from author's portal

```{r}
message("Loading Seurat object...")
data.seurat <- readRDS(input_rds)
message(ncol(data.seurat), " cells were loaded")
message(nrow(data.seurat), " genes were loaded")
```

### I.2. Check Seurat object

```{r, fig.width=8, fig.height=8}
# Visualize QC metrics as a violin plot
DimPlot(data.seurat, reduction = "umap", raster = TRUE, group.by = "batch")
```

```{r, fig.width=8, fig.height=8}
# Visualize QC metrics as a violin plot
DimPlot(data.seurat, reduction = "umap", raster = TRUE, group.by = "new_clusters3")
```

```{r, fig.width=8, fig.height=8}
# Visualize QC metrics as a violin plot
DimPlot(data.seurat, reduction = "umap", raster = TRUE, group.by = "sex")
```

### I.3. Reading global metadata from authors

```{r}
data.metadata <- fread(file = input_metadata, stringsAsFactors = F, data.table = F)
rownames(data.metadata) <- data.metadata$V1
data.metadata <- data.metadata[,-1]
data.metadata
```

Seems to match the Seurat object

## II. h5ad

NOTE: I could use sceasy or SeuratDisk option, but it does not do what I want. And there are many missing fields. So I'll create the h5ad from scratch.

# Prepare Cell attributes to put in h5ad

```{r}
# Cell metadata
obs <- data.seurat@meta.data
```

# Raw matrix (I remove colnames and rownames)

```{r}
# Raw count matrix
raw.X <- GetAssayData(object = data.seurat, assay = "RNA", layer = "counts")
```

# Prepare Gene metadata to put in h5ad

```{r}
# Gene metadata
var <- as.data.frame(GetAssay(data.seurat, assay = "RNA")@features)
var$Gene <- rownames(var)
#var$Accession <- gene.mapping[rownames(var),"Name"]
var
```

# Normalized matrix

```{r}
# Normalized count matrix
X <- GetAssayData(object = data.seurat, assay = "RNA", layer = "data")
```

# Embeddings

```{r}
obsm <- NULL
reductions <- names(data.seurat@reductions)
if (length(reductions) > 0) {
  obsm <- sapply(
    reductions,
    function(name) as.matrix(Seurat::Embeddings(data.seurat, reduction = name)),
    simplify = FALSE
  )
  names(obsm) <- paste0("X_", tolower(names(data.seurat@reductions))) # More compatible with usual format
}
```

# Global metadata in @misc

```{r}
uns <- data.seurat@misc
uns$default_embedding <- "X_umap"
```

# Create h5ad file

```{r}
# AnnData object creation using reticulate
adata <- ad$AnnData(X = Matrix::t(X), raw = list(X = Matrix::t(raw.X), var = var), obs = obs, var = var, obsm = obsm, uns = uns)
```

# Write object to file

```{r}
adata$write(output_h5ad, compression = "gzip")
```