---
title: "Emma Pan Neuro (Control + ND75KD)"
description: "Figure - DE genes between ND75KD and Control, by annotation"
author: "Vincent Gardeux"
date: "2025/02/04"
updated: "2025/02/04"
output:
  html_document:
    df_print: paged
---

  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/data/gardeux/Neuro_Droso_ND75KD/")
```

## Libraries & functions

First, I'm loading the required libraries & functions

```{r}
#suppressPackageStartupMessages(library(Seurat, lib.loc = "/software/Seuratv4/lib/")) # Seurat v4.4.0, library for single-cell analysis
#suppressPackageStartupMessages(library(SeuratObject, lib.loc = "/software/Seuratv4/lib/")) # For compatibility purpose
suppressPackageStartupMessages(library(Seurat)) # Seurat v5
suppressPackageStartupMessages(library(SeuratObject)) # Seurat v5
suppressPackageStartupMessages(library(data.table)) # For writing DE gene file
suppressPackageStartupMessages(library(ggplot2)) # For plotting
suppressPackageStartupMessages(library(ggpubr)) # For grid plotting
suppressPackageStartupMessages(library(ggrepel)) # For grid plotting
suppressPackageStartupMessages(library(crayon)) # Just for bolding the console output :D

cat(bold("Seurat"), "version", as.character(packageVersion("Seurat")), "\n")
cat(bold("SeuratObject"), "version", as.character(packageVersion("SeuratObject")), "\n")
cat(bold("data.table"), "version", as.character(packageVersion("data.table")), "\n")
cat(bold("ggplot2"), "version", as.character(packageVersion("ggplot2")), "\n")
cat(bold("ggpubr"), "version", as.character(packageVersion("ggpubr")), "\n")
cat(bold("ggrepel"), "version", as.character(packageVersion("ggrepel")), "\n")

# Color-blind friendly palette 3-colors
cbPalette <- c("#E69F00", "#000000", "#56B4E9")
steinPalette <- c("#80C980", "#BDADD4", "#376CB0", "#FBBF85", "#F0027E")
tissuePalette <- list(`Pan_neuro_control`="#E69F00", `Pan_neuro_ND75KD`="#56B4E9")
celltypePalette <- list(`unannotated`="#D3D3D3", `epithelial cell`="#E0AC69", `skeletal muscle of head`="#895129", `cone cell`="#FFFFB7", `outer photoreceptor cell`="#FFF060", `photoreceptor cell`="#F2DB00", `photoreceptor-like cell`="#F0C200", `adult optic chiasma glial cell`="#d896ff", `adult brain cell body glial cell`="#efbbff", `adult brain perineurial glial cell`="#BCBDDC", `adult lamina epithelial/marginal glial cell`="#9E9AC8", `adult reticular neuropil associated glial cell`="#756BB1", `optic lobe associated cortex glial cell`="#54278F", `cholinergic neuron`="#E41A1C", `gabaergic neuron`="#377EB8", `glutamatergic neuron`="#4DAF4A", `dopaminergic neuron`="#984EA3", `serotonergic neuron`="#FF7F00", `OPN neuron`="#A6CEE3", `columnar neuron T1`="#DEEBF7", `T neuron T3`="#9ECAE1", `T neuron T4/T5`="#3182BD", `distal medullary amacrine neuron Dm3`="#c6e6ba", `distal medullary amacrine neuron Dm8`="#A1D99B", `distal medullary amacrine neuron Dm9`="#31A354", `proximal medullary amacrine neuron Pm4`="#006D2C", `centrifugal neuron C2`="#7FC97F", `centrifugal neuron C3`="#BEAED4", `medullary intrinsic neuron Mi1`="#d8d5eb", `medullary intrinsic neuron Mi4`="#BCBDDC", `medullary intrinsic neuron Mi15`="#756BB1", `transmedullary neuron Tm1`="#b4dee9", `transmedullary neuron Tm2`="#bfdfe7", `transmedullary neuron Tm9`="#7eb5c4", `transmedullary neuron Tm20`="#9fcad5", `transmedullary Y neuron TmY14`="#5da1b3", `transmedullary Y neuron TmY5a`="#3c8ca2", `transmedullary Y neuron TmY8`="#1a7790", `tm5ab`="#014636", `lamina monopolar neuron L1 + L2`="#FDD3A0", `lamina monopolar neuron L3`="#FDBE85", `lamina monopolar neuron L4`="#FD8D3C", `lamina monopolar neuron L5`="#E6550D", `lamina wide-field 2 neuron`="#A63603", `lobular columnar neuron LC12`="#BC80BD", `olfactory receptor neuron`="#CCEBC5", `Poxn neuron`="#FFED6F", `hemocyte`="#999999", `kenyon cell`="#B3E2CD", `gamma kenyon cell`="#FDCDAC", `pigment cell`="#EE82EE")


# Random seed
set.seed(42)
```

## Parameters

```{r}
input_seurat_object <- "./data/Pan_neuro_integrated_FINAL.rds"
output_figure_prefix <- "./figures/Figure_1X/Volcano_plot_regulons"
SCENIC_regulons_path <- "./data/Pan_neuro_integrated_regulons_aucell.tsv"
marker_regulon_output <- "./data/Pan_neuro_integrated_markers_regulons"
```

## I. Reading Seurat object

First step is to read the Seurat objects, previously created

```{r}
message("Loading Seurat object...")
data.seurat <- readRDS(input_seurat_object)
message(ncol(data.seurat), " cells were loaded")
message(nrow(data.seurat), " genes were loaded")
```

## II. Differentially activated regulons (based on non-binarized data)

```{r}
# 1. Put SCENIC regulon in a new assay
data.scenic <- fread(SCENIC_regulons_path, data.table = F)
rownames(data.scenic) <- data.scenic$Cell
data.scenic <- data.scenic[,-1]
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\+\\)", replacement = "-activating-regulon")
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\-\\)", replacement = "-repressing-regulon")
colnames(data.scenic) <- paste0("SCENIC-", colnames(data.scenic))
data.seurat@assays[["SCENIC"]] <- CreateAssayObject(data = t(data.scenic)[,colnames(data.seurat)], check.matrix = F)

# 2. Run DE
data.seurat <- SetIdent(data.seurat, value = "seurat_clusters")
all.markers <- FindAllMarkers(data.seurat, assay = "SCENIC", slot = "data", verbose = T, test.use = "wilcox", random.seed = 42) # For unknown reasons Seuratv4 fails. So I use Seuratv5.

fwrite(all.markers, file = paste0(marker_regulon_output, ".tsv"), sep = "\t")
all.markers
```

### DE in cluster 2

sima (p.5) and crc (p.19) are there, but with low fold change: sima(log2FC = 0.7389760) & crc(log2FC = 0.3106056)

```{r}
all.markers.2 <- subset(all.markers, cluster == "2")
all.markers.2 <- all.markers.2[order(abs(all.markers.2$avg_log2FC), decreasing = T),]
all.markers.2
```

## Volcano Plot

```{r, fig.width=7, fig.height=5}
# Add required columns
all.markers.2$neg_log10_p_val_adj <- -log10(all.markers.2$p_val_adj)
all.markers.2$neg_log10_p_val_adj[is.infinite(all.markers.2$neg_log10_p_val_adj)] <- 300
all.markers.2$category <- "Non-significant"
all.markers.2$category[all.markers.2$p_val_adj <= 0.05 & all.markers.2$avg_log2FC >= 1] <- "Upregulated"
all.markers.2$category[all.markers.2$p_val_adj <= 0.05 & all.markers.2$avg_log2FC <= -1] <- "Downregulated"
all.markers.2$regulon <- gsub(x = all.markers.2$gene, pattern = "SCENIC-", replacement = "")
all.markers.2$regulon <- gsub(x = all.markers.2$regulon, pattern = "-activating-regulon", replacement = "")

# Choose coloring
colors.volcano <- c("Non-significant" = "grey", "Upregulated" = "red", "Downregulated" = "blue")

# Genes to annotate
genes_to_label <- c("sage", "Hr39", "Atf-2", "gt", "sima", "crc", "CG10348", "CG6808", "esg", "bowl", "NC2alpha", "su(Hw)", "Mtk", "ovo", "Myc", "odd", "Mabi", "CG4374", "term", "CG1602")

# Plot
p <- ggplot(all.markers.2, aes(x = avg_log2FC, y = neg_log10_p_val_adj, color = category)) +
  geom_point(alpha = 0.7, size = 3) +
  scale_color_manual(values = colors.volcano) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +
  labs(x = "Log2 Fold Change", y = "-Log10 Adjusted p-value", color = NULL) +
  theme_minimal() +
  geom_text_repel(data = subset(all.markers.2, regulon %in% genes_to_label), aes(label = regulon), size = 4, segment.color = "black", box.padding = 1, point.padding = 1, max.overlaps = 100)
p
```

```{r}
ggsave(p, filename = paste0(output_figure_prefix, ".pdf"), width = 7, height = 5, bg = "white")
ggsave(p, filename = paste0(output_figure_prefix, ".png"), width = 7, height = 5, dpi = 1000, bg = 'white')
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("crc", "SCENIC_crc_activating_regulon", "SCENIC_crc_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("sima", "SCENIC_sima_activating_regulon", "SCENIC_sima_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("su(Hw)", "SCENIC_su(Hw)_activating_regulon", "SCENIC_su(Hw)_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("gt", "SCENIC_gt_activating_regulon", "SCENIC_gt_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("sage", "SCENIC_sage_activating_regulon", "SCENIC_sage_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("Hr39", "SCENIC_Hr39_activating_regulon", "SCENIC_Hr39_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```

### DE in cluster 23

sima is on the first page, and crc is on the second page.

```{r}
all.markers.23 <- subset(all.markers, cluster == "23")
all.markers.23 <- all.markers.23[order(abs(all.markers.23$avg_log2FC), decreasing = T),]
all.markers.23
```

## Volcano Plot

```{r, fig.width=7, fig.height=5}
# Add required columns
all.markers.23$neg_log10_p_val_adj <- -log10(all.markers.23$p_val_adj)
all.markers.23$neg_log10_p_val_adj[is.infinite(all.markers.23$neg_log10_p_val_adj)] <- 300
all.markers.23$category <- "Non-significant"
all.markers.23$category[all.markers.23$p_val_adj <= 0.05 & all.markers.23$avg_log2FC >= 1] <- "Upregulated"
all.markers.23$category[all.markers.23$p_val_adj <= 0.05 & all.markers.23$avg_log2FC <= -1] <- "Downregulated"
all.markers.23$regulon <- gsub(x = all.markers.23$gene, pattern = "SCENIC-", replacement = "")
all.markers.23$regulon <- gsub(x = all.markers.23$regulon, pattern = "-activating-regulon", replacement = "")

# Choose coloring
colors.volcano <- c("Non-significant" = "grey", "Upregulated" = "red", "Downregulated" = "blue")

# Genes to annotate
genes_to_label <- c("CG5245", "CG9876", "CG33213", "p53", "Hey", "Myc", "dbr", "dany", "insv", "hkb", "Ssrp", "jumu", "hang", "gsb", "Hr39", "sage", "Atf-2", "gt", "sima", "crc", "CG6808", "esg", "bowl", "NC2alpha", "Myc", "odd", "term", "CG1602")

# Plot
p <- ggplot(all.markers.23, aes(x = avg_log2FC, y = neg_log10_p_val_adj, color = category)) +
  geom_point(alpha = 0.7, size = 3) +
  scale_color_manual(values = colors.volcano) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +
  labs(x = "Log2 Fold Change", y = "-Log10 Adjusted p-value", color = NULL) +
  theme_minimal() +
  geom_text_repel(data = subset(all.markers.23, regulon %in% genes_to_label), aes(label = regulon), size = 4, segment.color = "black", box.padding = 1, point.padding = 1, max.overlaps = 100)
p
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("Hey", "SCENIC_Hey_activating_regulon", "SCENIC_Hey_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("NC2alpha", "SCENIC_NC2alpha_activating_regulon", "SCENIC_NC2alpha_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("p53", "SCENIC_p53_activating_regulon", "SCENIC_p53_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```

```{r, fig.height = 5, fig.width = 15}
FeaturePlot(data.seurat, reduction = "SCENICbinarizedumap", features = c("bowl", "SCENIC_bowl_activating_regulon", "SCENIC_bowl_activating_regulon_binarized"), order = T, pt.size = 0.5, ncol = 3)
```


## VIII. Differentially activated regulons in Pan neuro CTRL (based on non-binarized data)

```{r}
# 1. Subset the object
Idents(data.seurat) <- "orig.ident"
data.seurat@assays[["SCENIC"]] <- NULL
data.seurat.ctrl <- subset(x = data.seurat, idents = c("Pan_neuro_control"))

# 2. Put SCENIC regulon in a new assay
data.scenic <- fread(SCENIC_regulons_path, data.table = F)
rownames(data.scenic) <- data.scenic$Cell
data.scenic <- data.scenic[,-1]
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\+\\)", replacement = "-activating-regulon")
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\-\\)", replacement = "-repressing-regulon")
colnames(data.scenic) <- paste0("SCENIC-", colnames(data.scenic))
data.seurat.ctrl@assays[["SCENIC"]] = CreateAssayObject(data = t(data.scenic)[,colnames(data.seurat.ctrl)], check.matrix = F)

# 3. Run DE
data.seurat.ctrl <- SetIdent(data.seurat.ctrl, value = "seurat_clusters")
all.markers <- FindAllMarkers(data.seurat.ctrl, assay = "SCENIC", slot = "data", verbose = T, test.use = "wilcox", random.seed = 42) # For unknown reasons Seuratv4 fails. So I use Seuratv5.

fwrite(all.markers, file = paste0(marker_regulon_output, "_ctrl.tsv"), sep = "\t")

all.markers
```

### DE in cluster 2

```{r}
ctrl.markers.2 <- subset(all.markers, cluster == "2")
ctrl.markers.2 <- ctrl.markers.2[order(abs(ctrl.markers.2$avg_log2FC), decreasing = T),]
ctrl.markers.2
```

Nothing is significant...
(because of the number of cells 4 CTRL vs 813 ND75KD)

### DE in cluster 23

```{r}
ctrl.markers.23 <- subset(all.markers, cluster == "23")
ctrl.markers.23 <- ctrl.markers.23[order(abs(ctrl.markers.23$avg_log2FC), decreasing = T),]
ctrl.markers.23
```

Was not run because cluster 23 has less than 3 cells for CTRL samples (2 CTRL cells vs 331 ND75KD cells)

## IX. Differentially activated regulons in Pan neuro ND75KD (based on non-binarized data)

```{r}
# 1. Subset the object
Idents(data.seurat) <- "orig.ident"
data.seurat@assays[["SCENIC"]] <- NULL
data.seurat.ctrl <- subset(x = data.seurat, idents = c("Pan_neuro_ND75KD"))

# 2. Put SCENIC regulon in a new assay
data.scenic <- fread(SCENIC_regulons_path, data.table = F)
rownames(data.scenic) <- data.scenic$Cell
data.scenic <- data.scenic[,-1]
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\+\\)", replacement = "-activating-regulon")
colnames(data.scenic) <- gsub(x = colnames(data.scenic), pattern = "\\(\\-\\)", replacement = "-repressing-regulon")
colnames(data.scenic) <- paste0("SCENIC-", colnames(data.scenic))
data.seurat.ctrl@assays[["SCENIC"]] = CreateAssayObject(data = t(data.scenic)[,colnames(data.seurat.ctrl)], check.matrix = F)

# 3. Run DE
data.seurat.ctrl <- SetIdent(data.seurat.ctrl, value = "seurat_clusters")
all.markers <- FindAllMarkers(data.seurat.ctrl, assay = "SCENIC", slot = "data", verbose = T, test.use = "wilcox", random.seed = 42) # For unknown reasons Seuratv4 fails. So I use Seuratv5.

fwrite(all.markers, file = paste0(marker_regulon_output, "_nd75kd.tsv"), sep = "\t")

all.markers
```

### DE in cluster 2

```{r}
nd75kd.markers.2 <- subset(all.markers, cluster == "2")
nd75kd.markers.2 <- nd75kd.markers.2[order(abs(nd75kd.markers.2$avg_log2FC), decreasing = T),]
nd75kd.markers.2
```

More or less same results than with all cells. There are a few less significant ones.
And 3 'new' downregulated ones: "CG14655", "nau", "eg"

## Volcano Plot

```{r, fig.width=7, fig.height=5}
# Add required columns
nd75kd.markers.2$neg_log10_p_val_adj <- -log10(nd75kd.markers.2$p_val_adj)
nd75kd.markers.2$neg_log10_p_val_adj[is.infinite(nd75kd.markers.2$neg_log10_p_val_adj)] <- 300
nd75kd.markers.2$category <- "Non-significant"
nd75kd.markers.2$category[nd75kd.markers.2$p_val_adj <= 0.05 & nd75kd.markers.2$avg_log2FC >= 1] <- "Upregulated"
nd75kd.markers.2$category[nd75kd.markers.2$p_val_adj <= 0.05 & nd75kd.markers.2$avg_log2FC <= -1] <- "Downregulated"
nd75kd.markers.2$regulon <- gsub(x = nd75kd.markers.2$gene, pattern = "SCENIC-", replacement = "")
nd75kd.markers.2$regulon <- gsub(x = nd75kd.markers.2$regulon, pattern = "-activating-regulon", replacement = "")

# Choose coloring
colors.volcano <- c("Non-significant" = "grey", "Upregulated" = "red", "Downregulated" = "blue")

# Genes to annotate
genes_to_label <- c("sage", "Hr39", "Atf-2", "gt", "sima", "crc", "CG10348", "esg", "bowl", "ovo", "Mabi", "CG4374", "term", "CG1602", "CG14655", "nau", "eg")

# Plot
p <- ggplot(nd75kd.markers.2, aes(x = avg_log2FC, y = neg_log10_p_val_adj, color = category)) +
  geom_point(alpha = 0.7, size = 3) +
  scale_color_manual(values = colors.volcano) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +
  labs(x = "Log2 Fold Change", y = "-Log10 Adjusted p-value", color = NULL) +
  theme_minimal() +
  geom_text_repel(data = subset(nd75kd.markers.2, regulon %in% genes_to_label), aes(label = regulon), size = 4, segment.color = "black", box.padding = 1, point.padding = 1, max.overlaps = 100)
p
```

### DE in cluster 23

```{r}
nd75kd.markers.23 <- subset(all.markers, cluster == "23")
nd75kd.markers.23 <- nd75kd.markers.23[order(abs(nd75kd.markers.23$avg_log2FC), decreasing = T),]
nd75kd.markers.23
```

More or less same results than with all cells, a few less significant ones.
One new one: "CG10348"

## Volcano Plot

```{r, fig.width=7, fig.height=5}
# Add required columns
nd75kd.markers.23$neg_log10_p_val_adj <- -log10(nd75kd.markers.23$p_val_adj)
nd75kd.markers.23$neg_log10_p_val_adj[is.infinite(nd75kd.markers.23$neg_log10_p_val_adj)] <- 300
nd75kd.markers.23$category <- "Non-significant"
nd75kd.markers.23$category[nd75kd.markers.23$p_val_adj <= 0.05 & nd75kd.markers.23$avg_log2FC >= 1] <- "Upregulated"
nd75kd.markers.23$category[nd75kd.markers.23$p_val_adj <= 0.05 & nd75kd.markers.23$avg_log2FC <= -1] <- "Downregulated"
nd75kd.markers.23$regulon <- gsub(x = nd75kd.markers.23$gene, pattern = "SCENIC-", replacement = "")
nd75kd.markers.23$regulon <- gsub(x = nd75kd.markers.23$regulon, pattern = "-activating-regulon", replacement = "")

# Choose coloring
colors.volcano <- c("Non-significant" = "grey", "Upregulated" = "red", "Downregulated" = "blue")

# Genes to annotate
genes_to_label <- c("CG5245", "CG9876", "CG33213", "p53", "Hey", "Myc", "dbr", "dany", "insv", "hkb", "gsb", "Hr39", "sage", "Atf-2", "gt", "sima", "crc", "CG6808", "esg", "bowl", "NC2alpha", "Myc", "odd", "term", "CG1602", "CG10348")

# Plot
p <- ggplot(nd75kd.markers.23, aes(x = avg_log2FC, y = neg_log10_p_val_adj, color = category)) +
  geom_point(alpha = 0.7, size = 3) +
  scale_color_manual(values = colors.volcano) +
  geom_vline(xintercept = c(-1, 1), linetype = "dashed", color = "black") +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black") +
  labs(x = "Log2 Fold Change", y = "-Log10 Adjusted p-value", color = NULL) +
  theme_minimal() +
  geom_text_repel(data = subset(nd75kd.markers.23, regulon %in% genes_to_label), aes(label = regulon), size = 4, segment.color = "black", box.padding = 1, point.padding = 1, max.overlaps = 100)
p
```
